\textbf{Computational Fuzzy Extractors}

### 2013-03-26 trunk/CompFuzzExt.tex
Working on wording of introduction

Authentication generally requires secrets.  Unfortunately, many useful  sources that contain sufficient entropy for a secret  are also noisy, and provide similar, but not identical secret values at each invocation (examples of such sources include biometrics~\cite{daugman2004}, human memory~\cite{zviran1993comparison}, pictorial passwords~\cite{brostoff2000passfaces}, measurements of capacitance~\cite{tuyls2006puf}, timing~\cite{suh2007physical}, motion~\cite{castelluccia2005shake},  quantum information~\cite{bennett1988privacy} etc.).  Information reconciliation protocols~\cite{bennett1988privacy} remove the noise without revealing the secrets.

One of the primary building blocks for authentications is reliable key derivation.  Fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} achieve reliable key derivation from noisy sources.  (\bnote{Are these references for fuzzy extractors or secure sketches?} as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).  The setting 
%We will specifically focus on a one-round information reconciliation  mechanism called ``secure sketch,''  which, aside from its immediate application to authentication, is also
%a crucial ingredient in the construction of fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} (as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).
consists of  two algorithms: Generate (used once) and Reproduce (used subsequently).  The Generate ($\gen$) algorithm takes an input $w$ and produces a key $r$ and a public value $p$.  This information allows
the Reproduce ($\rep$) algorithm to reproduce $r$ given $p$ and some value $w'$ that is close (according to some predefined metric, such as Hamming distance) to $w$. 
Crucially for security,  knowledge of $p$ should not reveal $r$; that is, $r$ should be uniformly distributed conditioned on $p$.  This feature is needed because $p$ is not secret: for example, in a single-user setting (where the user wants to recover the original reading $w$ from a subsequent reading $w'$), it would be stored in the clear; and in a key agreement application~\cite{Boyen05secureremote} (where two parties have $w$ and $w'$, respectively), it would be transmitted between the parties.

As defined in \cite{DBLP:journals/siamcomp/DodisORS08}, the entropy loss of a fuzzy extractor is the difference in the entropy of $w$ and the length of the key $r$.  A goal of fuzzy extractor constructions is to minimize the entropy loss, increasing the security of the resulting application.  The resulting secret key may be too short to be useful if the entropy loss is too high.  Traditionally, fuzzy extractors are defined as information-theoretic objects.  In the information-theoretic setting, some entropy loss is necessary as the value $p$ contains enough information to reproduce $r$ from any close value $w'$.

The main novel tool in constructing fuzzy extractors is the secure sketch.  The secure sketch produces a public value $s$ that allows recovery of $w$ from any close value $w'$.  The security requirement is that $w$ has high min-entropy conditioned on $s$.  The key $r$ is formed by applying a randomness extractor~\cite{nisan1993randomness} to $w$.
%Secret entropy of convenient sources is often at a premium; a goal of secure sketch constructions is to minimize the entropy loss, increasing the security of the resulting application. When used in a fuzzy extractor, the resulting secret key may be too short to be useful if the entropy loss is too high. 
The minimum entropy loss for secure sketches has been quantified in past work.  Even if we don't require perfect correctness (i.e., allow $\rep$ to err with small probability~\cite[Section 8]{DBLP:journals/siamcomp/DodisORS08}), the entropy loss of a secure sketch that can reconcile $t$ errors with high probability is bounded by the redundancy of
the best code that can correct $t$ random errors with high probability~\cite[Proposition 8.2]{DBLP:journals/siamcomp/DodisORS08}.  Known constructions of fuzzy extractors use a secure sketch suffer from this bound.
%(the argument is the same as in \cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}\lnote{probably we should double-check that this is correct}).

\paragraph {Our Contributions}
To see if entropy loss in fuzzy extractors can be improved, we explore fuzzy extractors and secure sketches with the relaxed secrecy requirement that is computational rather than information-theoretic.  Our overall goal is to produce a fuzzy extractor~(with small entropy loss), where the key $r$ is pseudorandom even conditioned on $p$.

@@

Another possible relaxation is to require that the value $w$ is still unpredictable conditioned on $s$.  Using an extractor with a reconstruction would still yield a pseudorandom key, see~\cite[Lemma 6]{DBLP:conf/eurocrypt/HsiaoLR07}.  We consider sketches that preserve the weaker notion of unpredictability for the Hamming metric.  In \thref{thm:imp of unp entropy}, we show  a significant loss in unpredictability is necessary for a secure sketch.  The entropy loss is proportional to the volume of the ball of radius $t$.

Both of these impossibility results arise because a sketch must function like an error-correcting code.  To avoid these impossibility results, we give up on reconstructing $w$ and focus directly on the entropy loss in fuzzy extractors.  
%to focus on producing a consistent string $\recout$  each time $\rec(w', s)$ is called.  This string $\recout$ is  required to have high HILL entropy even conditioned on $s$. Because this object \emph{conducts} the entropy from $w$ to $\recout$ and tolerates noise in its input, following \cite{CRVW02,KanukurthiR09}, we call it a \emph{computational fuzzy conductor}.
We build \emph{lossless} computational fuzzy extractors, in which $r$ is pseudorandom conditioned on $p$ and $r$ is as long as the entropy of $w$.  % without conditioning on anything---i.e., before $\gen$ is run (see~\thref{thm:lossless secure conductor log})).  
We demonstrate how to build such lossless fuzzy extractors~(\consref{cons:informal construction}) based on the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey},  which says that decoding a random linear code is computationally difficult.

%We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
%We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
%\lnote{it would be good to find another application}

Our lossless fuzzy extractor is built via the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} already used in prior work.  To be able to apply the LWE assumption, we use a random linear code (unfortunately, our decoding algorithm can only reconcile a logarithmic number of differences).  We utilize the recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} that shows security of LWE when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval.  The LWE error vector, in our case, is the input $w$ itself.  To utilize the result of~\cite{dottling2012}, we need to assume that the input comes from the uniform distribution. We are able to relax this limitation somewhat and allow $w$ to come from a symbol-fixing source~\cite{KZ07} (where each dimension is either uniform or fixed). This relaxation requires new results about the hardness of LWE when samples have a fixed~(and adversarially known) error vector, which may be of independent interest~(\thref{thm:blockLWE}).

@@

\subsection{Proof of \thref{thm:impSketchArbitraryW}}
\label{sec:proof of thm sketch implies code}
\begin{proof}
  Let $W$ be an arbitrary distribution of min-entropy $m$.  Let $Y_s$ be a collection of distributions over $\mathcal{M}$ giving rise to $Y$ such that $\Hav(Y | \sketch(W))\geq k$ such that 
\[ 
\delta^{\mathcal{D}_{s_{sec}}}((W, \sketch(W)), (Y, \sketch(W)))<\epsilon.
\]  
where  $s_{sec} = O(t(s_{neigh}+s_{rec}))$.  One such $Y$ must exist by the definition of conditional HILL entropy.
%Let $X$ be arbitrarily but independently distributed over $\mathcal{M}$. Let $\delta^D((X, \sketch(X)), (W, \sketch(X)))<\epsilon$ for negligible $\epsilon$ and $D$ of size at least $s=O(s_{neigh}+s_{rec})$.  This implies that $H^{\hill}_{\epsilon, s}(X|\sketch(X))\geq k$.  
Define $D$ as:
\begin{enumerate}
\item Input $z\in\mathcal{M}, s \in\{0, 1\}^*, t$.
\item For all $1\leq t'\leq t$: 
\subitem  $z'\leftarrow \sample(z, t')$.
\subitem If $\rec(z', s) \neq  z$ output $0$.
\item Output $1$.
\end{enumerate}
  By correctness of the sketch $ \Pr[D(W, \sketch(W)) =1]\ge 1-t\delta$.  Since 
$\delta^D((Y, \sketch(W)), (Y, \sketch(W)))<\epsilon.$ then $\Pr[D(Y, SS(W)) = 1]>1-\epsilon-t\delta$.  By Markov inequality there exists a set $S_W$ such that $\Pr[S_W]\geq 1/2$ and $\Pr[ D(Y, s) =1 | \sketch(W) = s \wedge s\in S_W]> 1- 2(\epsilon + t\delta)$.  Consider the random variable $1_{S_W}$ that is $1$ when $s\in S_X$ and otherwise.  Then \bnote{I am splitting average min-entropy based on an indicator variable, everyone is okay with this?}
\[
\Hav(Y  | W) = -\log \left( \Pr[W\in 1_{S_W}] 2^{-\Hav(Y | W\in 1_{S_W})} + \Pr[W\not\in 1_{S_W}] 2^{-\Hav (Y|W\not\in 1_{S_W})} \right)
\] 
Rearranging terms:
\begin{align*}
\Pr[W\in 1_{S_W}] 2^{-\Hav(Y | W\in 1_{S_W})} &= 2^{-\Hav(Y  | W) } - \Pr[W\not\in 1_{S_W}] 2^{-\Hav (Y|W\not\in 1_{S_W})}\\
&=2^{-\Hav(Y  | W) } = 2^{-k}
\end{align*}
This means that $2^{-\Hav(Y | W\in 1_{S_W})} \geq 2^{-k}/\Pr[W\in 1_{S_W}] = 2^{-k+1}$.  That is, $\Hav(Y|W\in 1_{S_W})\geq k-1$.  Fix one value $s \in 1_{S_W}$ such that $\Hav(Y | s)\geq k-1$.
%This means at least one $x'$ where this equation holds.  Fix one such $x'$.  
%This implies that $D(W, \sketch(x'))\geq 1-\epsilon$.  
Thus, for all $t', 1\leq t'\leq t$ on a random $y$
\[ \Pr_{y\leftarrow Y_s}[y'\leftarrow \sample(y, t') \wedge \rec(y',s) = y]\geq 1-2(\epsilon+t\delta).\]  
%Thus, $\rec(\cdot, \sketch(x'))$ is a decoding procedure that succeeds on a random neighbor of a random codeword with probability all but $\epsilon$.  
Thus,  $Y$ is a $(t, 2(\epsilon+t\delta))$-average error Shannon code with recovery $\rec(\cdot,s)$.  The statement of the theorem follows by application of \lemref{lem:averageToMaximalError}.  
\end{proof}

>>>>

Authentication generally require a secret drawn from some high-entropy source.  One of the primary building blocks for authentication is reliable key derivation.  Unfortunately, many sources that contain sufficient entropy to derive a key are  noisy, and provide similar, but not identical secret values at each reading (examples of such sources include biometrics~\cite{daugman2004}, human memory~\cite{zviran1993comparison}, pictorial passwords~\cite{brostoff2000passfaces}, measurements of capacitance~\cite{tuyls2006puf}, timing~\cite{suh2007physical}, motion~\cite{castelluccia2005shake},  quantum information~\cite{bennett1988privacy} etc.).  %Information reconciliation protocols~\cite{bennett1988privacy} remove the noise without revealing the secrets.

Fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} achieve reliable key derivation from noisy sources.  (\bnote{Are these references for fuzzy extractors or secure sketches?} as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).  The setting 
%We will specifically focus on a one-round information reconciliation  mechanism called ``secure sketch,''  which, aside from its immediate application to authentication, is also
%a crucial ingredient in the construction of fuzzy extractors~\cite{DBLP:journals/siamcomp/DodisORS08} (as well as in other contexts, e.g.,~\cite{Boyen05secureremote,dodisWichs2009,chandran2010privacy}).
consists of  two algorithms: Generate (used once) and Reproduce (used subsequently).  The Generate ($\gen$) algorithm takes an input $w$ and produces a key $r$ and a public value $p$.  This information allows
the Reproduce ($\rep$) algorithm to reproduce $r$ given $p$ and some value $w'$ that is close to $w$ (according to some predefined metric, such as Hamming distance). 
Crucially for security,  knowledge of $p$ should not reveal $r$; that is, $r$ should be uniformly distributed conditioned on $p$.  This feature is needed because $p$ is not secret: for example, in a single-user setting (where the user wants to reproduce the key $r$ from a subsequent reading $w'$), it would be stored in the clear; and in a key agreement application~\cite{Boyen05secureremote} (where two parties have $w$ and $w'$, respectively), it would be transmitted between the parties.

%As defined in \cite{DBLP:journals/siamcomp/DodisORS08}, 
The entropy loss of a fuzzy extractor is the difference between the entropy of $w$ and the length of the derived key $r$.  A goal of fuzzy extractor constructions is to minimize the entropy loss, increasing the security of the resulting application.  The resulting secret key may be too short to be useful if the entropy loss is too high.  Traditionally, fuzzy extractors are defined as information-theoretic objects.  In the information-theoretic setting, some entropy loss is necessary as the value $p$ contains enough information to reproduce $r$ from any close value $w'$.

The main novel tool in constructing fuzzy extractors is the secure sketch.  The secure sketch produces a public value $s$ that allows recovery of $w$ from any close value $w'$.  The security requirement is that $w$ has high min-entropy conditioned on $s$.  The key $r$ is formed by applying a randomness extractor~\cite{nisan1993randomness} to $w$.
%Secret entropy of convenient sources is often at a premium; a goal of secure sketch constructions is to minimize the entropy loss, increasing the security of the resulting application. When used in a fuzzy extractor, the resulting secret key may be too short to be useful if the entropy loss is too high. 
The minimum entropy loss for secure sketches has been quantified in past work.  Even if $\rep$ is allowed to produce an incorrect value with small probability~\cite[Section 8]{DBLP:journals/siamcomp/DodisORS08}, 
%Even if we don't require perfect correctness (i.e., allow $\rep$ to err with small probability~\cite[Section 8]{DBLP:journals/siamcomp/DodisORS08}), 
the entropy loss of a secure sketch that can reconcile $t$ errors
 %with high probability 
 is bounded by the redundancy of
the best code that can correct $t$ random errors with high probability~\cite[Proposition 8.2]{DBLP:journals/siamcomp/DodisORS08}.  Constructions of fuzzy extractors that use a secure sketch inherit this bound.
%(the argument is the same as in \cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}\lnote{probably we should double-check that this is correct}).

\paragraph {Our Contributions}
To see if entropy loss in fuzzy extractors can be improved, we explore fuzzy extractors and secure sketches with a secrecy requirement that is computational rather than information-theoretic.  Our goal is to decrease the entropy loss in a fuzzy extractor by allowing the key $r$ to be pseudorandom conditioned on $p$.

@@

Another possible relaxation is to require that the value $w$ is unpredictable conditioned on $s$.  Using an extractor with a reconstruction would yield a pseudorandom key, see~\cite[Lemma 6]{DBLP:conf/eurocrypt/HsiaoLR07}.  We show lower bounds on the unpredictability loss of sketches in the Hamming metric.  We prove in \thref{thm:imp of unp entropy} that the unpredictability loss must be at least the volume of the ball of radius $t$.

Both of these impossibility results arise because a secure sketch functions like a decoder of an error-correcting code.  To avoid these impossibility results, we give up on building secure sketches and focus directly on the entropy loss in fuzzy extractors.

%to focus on producing a consistent string $\recout$  each time $\rec(w', s)$ is called.  This string $\recout$ is  required to have high HILL entropy even conditioned on $s$. Because this object \emph{conducts} the entropy from $w$ to $\recout$ and tolerates noise in its input, following \cite{CRVW02,KanukurthiR09}, we call it a \emph{computational fuzzy conductor}.
We build \emph{lossless} computational fuzzy extractors, where the derived key $r$ is as long as the entropy of the source $w$\footnote{If a source has enough entropy, the entropy losses of a information-theoretic fuzzy extractor can be overcome by running a pseudorandom generator on the derived key $r$.  However, when the source is low entropy, these entropy losses may not leave sufficient entropy to run a pseudorandom generator.}. % without conditioning on anything---i.e., before $\gen$ is run (see~\thref{thm:lossless secure conductor log})).  
Our construction of lossless fuzzy extractors~(\consref{cons:informal construction}) is based on the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey},  which says that decoding a random linear code is computationally difficult.  
%We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
%We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
%\lnote{it would be good to find another application} 
Our lossless fuzzy extractor is built via the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} used in prior work.  To apply the LWE assumption, we correct errors of a random linear code which limits our decoding algorithm to reconciling a logarithmic number of differences.

The recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} shows the hardness of LWE when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval.  This allows us to directly use our source $w$ as the LWE error vector.  To utilize the result of~\cite{dottling2012}, we need to assume that the input comes from the uniform distribution. We are able to relax this limitation and allow $w$ to come from a symbol-fixing source~\cite{KZ07} (each dimension is either uniform or fixed). This relaxation requires new results about the hardness of LWE when samples have a fixed~(and adversarially known) error vector, which may be of independent interest~(\thref{thm:blockLWE}).

@@

\subsection{Proof of \thref{thm:impSketchArbitraryW}}
\label{sec:proof of thm sketch implies code}
\begin{proof}
  Let $W$ be an arbitrary distribution of min-entropy $m$.  Let $Y_s$ be a collection of distributions over $\mathcal{M}$ giving rise to $Y$ such that $\Hav(Y | \sketch(W))\geq k$ such that 
\[ 
\delta^{\mathcal{D}_{s_{sec}}}((W, \sketch(W)), (Y, \sketch(W)))<\epsilon.
\]  
where  $s_{sec} = O(t(s_{neigh}+s_{rec}))$.  One such $Y$ must exist by the definition of conditional HILL entropy.
%Let $X$ be arbitrarily but independently distributed over $\mathcal{M}$. Let $\delta^D((X, \sketch(X)), (W, \sketch(X)))<\epsilon$ for negligible $\epsilon$ and $D$ of size at least $s=O(s_{neigh}+s_{rec})$.  This implies that $H^{\hill}_{\epsilon, s}(X|\sketch(X))\geq k$.  
Define $D$ as:
\begin{enumerate}
\item Input $z\in\mathcal{M}, s \in\{0, 1\}^*, t$.
\item For all $1\leq t'\leq t$: 
\subitem  $z'\leftarrow \sample(z, t')$.
\subitem If $\rec(z', s) \neq  z$ output $0$.
\item Output $1$.
\end{enumerate}
  By correctness of the sketch $ \Pr[D(W, \sketch(W)) =1]\ge 1-t\delta$.  Since 
$\delta^D((Y, \sketch(W)), (Y, \sketch(W)))<\epsilon.$ then $\Pr[D(Y, SS(W)) = 1]>1-\epsilon-t\delta$.  By Markov inequality there exists a set $S_W$ such that $\Pr[S_W]\geq 1/2$ and $\Pr[ D(Y, s) =1 | \sketch(W) = s \wedge s\in S_W]> 1- 2(\epsilon + t\delta)$.  Consider the random variable $1_{S_W}$ that is $1$ when $s\in S_X$ and otherwise.  Then,
\[
\Hav(Y  | W) = -\log \left( \Pr[W\in 1_{S_W}] 2^{-\Hav(Y | W\in 1_{S_W})} + \Pr[W\not\in 1_{S_W}] 2^{-\Hav (Y|W\not\in 1_{S_W})} \right)
\] 
Rearranging terms:
\begin{align*}
\Pr[W\in 1_{S_W}] 2^{-\Hav(Y | W\in 1_{S_W})} &= 2^{-\Hav(Y  | W) } - \Pr[W\not\in 1_{S_W}] 2^{-\Hav (Y|W\not\in 1_{S_W})}\\
&=2^{-\Hav(Y  | W) } = 2^{-k}
\end{align*}
This means that $2^{-\Hav(Y | W\in 1_{S_W})} \geq 2^{-k}/\Pr[W\in 1_{S_W}] = 2^{-k+1}$.  That is, $\Hav(Y|W\in 1_{S_W})\geq k-1$.  Fix one value $s \in 1_{S_W}$ such that $\Hav(Y | s)\geq k-1$.
%This means at least one $x'$ where this equation holds.  Fix one such $x'$.  
%This implies that $D(W, \sketch(x'))\geq 1-\epsilon$.  
Thus, for all $t', 1\leq t'\leq t$ on a random $y$
\[ \Pr_{y\leftarrow Y_s}[y'\leftarrow \sample(y, t') \wedge \rec(y',s) = y]\geq 1-2(\epsilon+t\delta).\]  
%Thus, $\rec(\cdot, \sketch(x'))$ is a decoding procedure that succeeds on a random neighbor of a random codeword with probability all but $\epsilon$.  
Thus,  $Y$ is a $(t, 2(\epsilon+t\delta))$-average error Shannon code with recovery $\rec(\cdot,s)$.  The statement of the theorem follows by application of \lemref{lem:averageToMaximalError}.  
\end{proof}
